---
title: "Adding new variables"
format:
  html:
    embed-resources: true
project:
  output-dir: results
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  message: false
---

## Loading packages

```{r}
#| label: load_packages_3

library("tidyverse")

source(file = here::here("R/99_proj_func.R"))
```

## Loading data files

```{r}
#| label: load_data_3

data_clean <- read_csv(file = here::here("data/data_clean.csv"))

PCA_clean <- read_csv(file = here::here("data/PCA_clean.csv"))
```

## Quantifying dataset diversity

```{r}
summary_counts <- data_clean |>
  summarise(
    n_CDs = n_distinct(CD),
    n_celltypes = n_distinct(cell_type)
  )

summary_counts
```

In this step, two summary variables are created to describe the structure of the dataset. Using the `summarise()` function, we count the number of unique CD markers and cell types present in the data. The result shows that the dataset includes 115 different CD markers and 38 distinct cell types, which highlights the biological diversity captured across tissues. This information provides a quick overview of how broad the dataset is and confirms that it contains enough variation to study immune cell development and activation patterns.

## Add variables

Here we enrich the dataset with two new variables:

-   hierarchy_level: assigns a numeric level (1–4) describing how specific or mature each cell type is.

-   lineage: groups cells into larger categories such as thymocytes, CD4 T cells, CD8 T cells, T cells, or B cells.

These new variables make it easier to visualize and compare immune-cell maturation patterns across tissues.

```{r}
#| label: add_var_3

level_3 <- c("Tgd", 
             "TCD8", 
             "TCD4", 
             "Bnaive", 
             "BnatEff", 
             "BIgM", 
             "BswMem",
             "Bdn", 
             "B27high", 
             "BnaiveTo", 
             "CC", 
             "CB", 
             "UnswtMem", 
             "SwtMem", 
             "PC")

data_aug <- data_clean |>
  mutate(hierarchy = case_when(cell_type == "Lymphs" ~ 1,
                               cell_type %in% c("T", "B") ~ 2,
                               cell_type %in% level_3 ~ 3,
                               .default = 4),
         .after = cell_type) |>
  add_lineage()
```

The same lineage classification is applied to the PCA dataset so that both data_aug and PCA_aug share identical labels.

```{r}
#| label: add_var_PCA_3

PCA_aug <- PCA_clean |>
  add_lineage()
```

## Writing data files

```{r}
#| label: write_data_3

write_csv(x = data_aug,
          file = here::here("data/data_aug.csv"))

write_csv(x = PCA_aug,
          file = here::here("data/PCA_aug.csv"))
```
