---
title: "Analysis 4"
format:
  html:
    embed-resources: true
project:
  output-dir: results
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  message: false
---

## Loading packages

```{r}
#| label: load_packages_8

library("tidyverse")
library("broom")
library("patchwork")

source(here::here(file = "R/99_proj_func.R"))
```

## Loading data files

```{r}
#| label: load_data_8

data_aug <- read_csv(here::here(file = "data/data_aug.csv"))
```

## Maturation of immune cells

This analysis explores how immune cells mature in different tissues (blood, tonsil, and thymus). The code groups and analyzes the data to investigate how the expression of several CDs changes depending on the type of cell and where it comes from.

To study these differences in a structured way, a linear model is applied to the log-transformed median fluorescence values (`log(MedQb)`). This model estimates how marker expression changes across cell types while accounting for variability between tissues and lineages. In this model, the naive cell population acts as the reference group, meaning that all other cell types are compared to it. This allows us to see how marker expression increases or decreases relative to naive cells as they progress through maturation or activation stages.

The data is first filtered based on `hierarchy` and `cell_type` as only the most specific cell subsets should be included in the analysis. Then the data is grouped and the data is filtered to ensure that there are more than one observation per `cell_type` in each group. Then the linear model applied. Each model produces an estimate of the average expression for each `cell_type` and its corresponding confidence interval. Lastly, the full estimates are calculated by retrieving and adding the `intercept` for each group. As the variable `cell_type` is lost during the transformation, it is recreated by using the variable `term`.

```{r}
#| label: apply_lm_8

columns <- c("CD", 
             "tissue",
             "lineage",
             "cell_type",
             "estimate",
             "conf.low", 
             "conf.high",
             "p.value",
             "is_significant",
             "any_significant")

data_analysis4 <- data_aug |>
  cell_type_order() |>
  filter(hierarchy %in% c(3, 4),
         !cell_type %in% c("Tgd", "TCD4", "TCD8")) |>
  select(-c("CVQb", "PEpos")) |>
  drop_na(MedQb) |>
  group_by(tissue, CD, lineage) |>
  filter(n() > n_distinct(cell_type)) |>
  fit_linear_model(log(MedQb) ~ cell_type)|>
  mutate(is_significant = case_when(term == "(Intercept)" ~ "reference",
                                    p.value < 0.05 ~ "yes",
                                    p.value >= 0.05 ~ "no"),
          any_significant = case_when(any(is_significant == "yes") ~ "yes",
                                     .default = "no")) |>
  mutate(
    intercept = estimate[term == "(Intercept)"],
    estimate = case_when(
      term == "(Intercept)" ~ estimate,
      term != "(Intercept)" ~ estimate + intercept),
    conf.low = case_when(
      term == "(Intercept)" ~ conf.low,
      term != "(Intercept)" ~ conf.low + intercept),
    conf.high = case_when(
      term == "(Intercept)" ~ conf.high,
      term != "(Intercept)" ~ conf.high + intercept),
    cell_type = case_when(
      term == "(Intercept)" & tissue == "blood" & lineage == "B cells" ~ "Bnaive",
      term == "(Intercept)" & tissue == "tonsil" ~ "BnaiveTo",
      term == "(Intercept)" & lineage == "CD4 T cells" ~ "TCD4naive", 
      term == "(Intercept)" & lineage == "CD8 T cells" ~ "TCD8naive",
      term == "(Intercept)" & tissue == "thymus"  ~ "DN34p",
    .default = str_remove(term, "cell_type"))) |>
  select(all_of(columns)) |>
  ungroup()
```

## Significant CDs

This section shows how many CD markers were significant (p \< 0.05) in each lineage.

```{r}
#| label: count_sig_CDs_8

cds_significant <- data_analysis4 |>
  filter(is_significant == "yes") |>      
  group_by(lineage) |>                    
  summarise(n_significant_CDs = n_distinct(CD)) |>
  arrange(desc(n_significant_CDs))        

cds_significant
```

A column plot of the count of significant CDs for each lineage is created below.

```{r}
#| label: plot_sig_CDs_8

sig_CD_count <- data_analysis4 |>
  plot_sig_CDs(by_variable = lineage, 
                 fill_color = "#fa2840")


sig_CD_count
```

The results are then combined into a single table called data_analysis4, which serves as the base for visualization and further interpretation. The function plot_cd_expression() is used to visualize these model results, plotting the estimated expression levels and their confidence intervals for each CD marker across tissues and cell types. The color indicates statistical significance, and the point shape represents the tissue of origin, allowing clear visual comparison between blood, tonsil, and thymus.

For the B cells, the analysis focuses on markers like CD11a, CD69, CD79b, and CD80, which are related to how these cells activate and differentiate. For T cells, the analysis is divided between CD4 and CD8 cells. The CD4 group includes markers like CD10, CD25, CD4_MEM-241, and CD62L, while the CD8 group includes CD27, CD38, CD71, and CD95. The plots compare expression levels across tissues, helping visualize how these cells evolve from early to more mature stages.

## B cells

```{r}
#| label: B_cells_8

cds_interest_b <- c("CD11a", "CD69", "CD79b", "CD80")

p1 <- data_analysis4 |>
  plot_cd_expression(lineage_filter = "B cells",
                     cds = cds_interest_b,
                     title = "Significant CDs during B cell maturation")

p1
```

**B cell markers:** CD11a (cell adhesion), CD69 (early activation), CD79b (B cell receptor signaling), CD80 (co-stimulation).

## T cells

```{r}
#| label: CD4_T_cells_8

cds_interest_cd4 <- c("CD10", "CD25", "CD4_MEM-241", "CD62L")

data_analysis4 |>
  plot_cd_expression(lineage_filter = c("Thymocytes", "CD4 T cells"),
                     cds = cds_interest_cd4,
                     cell_type_exclude = "CD8",
                     title = "Significant CDs during CD8 T cell maturation")
```

**CD4 T cell markers:** CD10 (early maturation), CD25 (activation), CD4_MEM-241 (memory), CD62L (migration).

```{r}
#| label: CD8_T_cells_8

cds_interest_cd8 <- c("CD27", "CD38", "CD71", "CD95")

data_analysis4 |>
  plot_cd_expression(lineage_filter = c("Thymocytes", "CD8 T cells"),
                     cds = cds_interest_cd8,
                     cell_type_exclude = "CD4",
                     title = "Significant CDs during CD8 T cell maturation")
```

**CD8 T cell markers:** CD27 (memory), CD38 (activation), CD71 (proliferation), CD95 (apoptosis).

Overall, the goal is to get a clearer picture of how immune cells change as they mature, how their protein expression shifts depending on the tissue, and how these differences relate to the naive reference cells. The comparison against naive populations makes it easier to see which markers increase or decrease as cells become more mature or activated. Using linear modeling together with visualization provides a clear and intuitive way to interpret these changes, helping to understand the development and activation of B and T cells across different tissues.

## Export plots

```{r}
#| label: export_plots_8

final_plot <- sig_CD_count / p1

save_plot(plot = final_plot, 
          filename = "08_key_plot.png",
          width = 7,
          height = 8.5)
```
