---
title: "Analysis 4"
format:
  html:
    embed-resources: true
project:
  output-dir: results
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  message: false
---

## Loading packages

```{r}
#| label: load_packages_8

library("tidyverse")
library("broom")
library("patchwork")

source(here::here(file = "R/99_proj_func.R"))
```

## Loading data files

```{r}
#| label: load_data_8

data_aug <- read_csv(here::here(file = "data/data_aug.csv"))
```

The aim of this analysis is to explore how different CD markers change across immune cell types and maturation stages in several tissues, using linear models and simple visualizations to highlight meaningful differences.

## Maturation of immune cells

This part prepares the data and runs the linear models. We keep only the detailed cell-type levels we care about and remove some unwanted columns and populations. After cleaning, we group the data by tissue, CD marker, and lineage, and then fit a linear model (`log(MedQb) ~ cell_type`) for each group.\
Once the models are fit, we classify whether each term is significant, rebuild the cell-type names, and adjust estimates by adding the intercept back in so everything stays on the same scale.

```{r}
#| label: apply_lm_8

columns <- c("CD", 
             "tissue",
             "lineage",
             "cell_type",
             "estimate",
             "conf.low", 
             "conf.high",
             "p.value",
             "is_significant",
             "any_significant")

data_analysis4 <- data_aug |>
  cell_type_order() |>
  filter(hierarchy %in% c(3, 4),
         !cell_type %in% c("Tgd", "TCD4", "TCD8")) |>
  select(-c("CVQb", "PEpos")) |>
  drop_na(MedQb) |>
  group_by(tissue, CD, lineage) |>
  filter(n() > n_distinct(cell_type)) |>
  fit_linear_model(log(MedQb) ~ cell_type)|>
  mutate(is_significant = case_when(term == "(Intercept)" ~ "reference",
                                    p.value < 0.05 ~ "yes",
                                    p.value >= 0.05 ~ "no"),
          any_significant = case_when(any(is_significant == "yes") ~ "yes",
                                     .default = "no")) |>
  mutate(
    intercept = estimate[term == "(Intercept)"],
    estimate = case_when(
      term == "(Intercept)" ~ estimate,
      term != "(Intercept)" ~ estimate + intercept),
    conf.low = case_when(
      term == "(Intercept)" ~ conf.low,
      term != "(Intercept)" ~ conf.low + intercept),
    conf.high = case_when(
      term == "(Intercept)" ~ conf.high,
      term != "(Intercept)" ~ conf.high + intercept),
    cell_type = case_when(
      term == "(Intercept)" & tissue == "blood" & lineage == "B cells" ~ "Bnaive",
      term == "(Intercept)" & tissue == "tonsil" ~ "BnaiveTo",
      term == "(Intercept)" & lineage == "CD4 T cells" ~ "TCD4naive", 
      term == "(Intercept)" & lineage == "CD8 T cells" ~ "TCD8naive",
      term == "(Intercept)" & tissue == "thymus"  ~ "DN34p",
    .default = str_remove(term, "cell_type"))) |>
  select(all_of(columns)) |>
  ungroup()
```

## Significant CDs

Here we check which CD markers actually show significant differences between cell types.\
We filter for terms with p \< 0.05, count how many CDs are significant within each lineage.

```{r}
#| label: count_sig_CDs_8

cds_significant <- data_analysis4 |>
  filter(is_significant == "yes") |>      
  group_by(lineage) |>                    
  summarise(n_significant_CDs = n_distinct(CD)) |>
  arrange(desc(n_significant_CDs))        

cds_significant
```

A column plot of the count of significant CDs for each lineage is created below.

```{r}
#| label: plot_sig_CDs_8

sig_CD_count <- data_analysis4 |>
  plot_sig_CDs(by_variable = lineage, 
                 fill_color = "#fa2840")


sig_CD_count
```

The results are then combined into a single table called data_analysis4, which serves as the base for visualization and further interpretation. The function `plot_cd_expression()` is used to visualize these model results, plotting the estimated expression levels and their confidence intervals for each CD marker across tissues and cell types. The color indicates statistical significance, and the point shape represents the tissue of origin, allowing clear visual comparison between blood, tonsil, and thymus.

For the B cells, the analysis focuses on markers like CD11a, CD69, CD79b, and CD80, which are related to how these cells activate and differentiate. For T cells, the analysis is divided between CD4 and CD8 cells. The CD4 group includes markers like CD10, CD25, CD4_MEM-241, and CD62L, while the CD8 group includes CD27, CD38, CD71, and CD95. The plots compare expression levels across tissues, helping visualize how these cells evolve from early to more mature stages.

## B cells

```{r}
#| label: B_cells_8

cds_interest_b <- c("CD11a", "CD69", "CD79b", "CD80")

p1 <- data_analysis4 |>
  plot_cd_expression(lineage_filter = "B cells",
                     cds = cds_interest_b,
                     subtitle = "B cells")

p1
```

**B cell markers:** CD11a (cell adhesion), CD69 (early activation), CD79b (B cell receptor signaling), CD80 (co-stimulation).

## T cells

```{r}
#| label: CD4_T_cells_8

cds_interest_cd4 <- c("CD10", "CD25", "CD4_MEM-241", "CD62L")

p2 <- data_analysis4 |>
  plot_cd_expression(lineage_filter = c("Thymocytes", "CD4 T cells"),
                     cds = cds_interest_cd4,
                     cell_type_exclude = "CD8",
                     subtitle = "CD4 T cells")

p2
```

**CD4 T cell markers:** CD10 (early maturation), CD25 (activation), CD4_MEM-241 (memory), CD62L (migration).

```{r}
#| label: CD8_T_cells_8

cds_interest_cd8 <- c("CD27", "CD38", "CD71", "CD95")

p3 <- data_analysis4 |>
  plot_cd_expression(lineage_filter = c("Thymocytes", "CD8 T cells"),
                     cds = cds_interest_cd8,
                     cell_type_exclude = "CD4",
                     subtitle = "CD8 T cells")

p3
```

**CD8 T cell markers:** CD27 (memory), CD38 (activation), CD71 (proliferation), CD95 (apoptosis).

Overall, the goal is to get a clearer picture of how immune cells change as they mature, how their protein expression shifts depending on the tissue, and how these differences relate to the naive reference cells. The comparison against naive populations makes it easier to see which markers increase or decrease as cells become more mature or activated. Using linear modeling together with visualization provides a clear and intuitive way to interpret these changes, helping to understand the development and activation of B and T cells across different tissues.

## Export plots

```{r}
#| label: export_plots_8

save_plot(plot = sig_CD_count, 
          filename = "08_key_plot_1.png")

p1 <- p1 + 
  labs(title = "Selected significant CDs during maturation stratified on lineage")

final_plot <- p1 + p2 + p3

save_plot(plot = final_plot, 
          filename = "08_key_plot_2.png",
          width = 22,
          height = 9)
```
